\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-17 arch.tex
europ+concepts touch ups with Javier, fig for arch and updates to arch

A prime motivation behind the \rx architecture was to design a
controller which could bring automated planning closer to low-level
control of the vehicle while maintaining its reactivity. Traditionally
planning has been deemed computationally expensive with planner
performance restricting robot reactivity
\cite{ghallab04,Dias:2003ua}. In control architectures where task
planning is embedded, planning consequently remains at an abstract
level with the assumption that it would otherwise impede system
reactiveness and that the environment can change at a faster rate than
the planner can plan for it. In such a situation, the agent may thrash
if the internal state of the plan gets out of synch with the actual
state of the world. As a result the planning problem managed {\em in
  situ} remains fairly detached from low-level control. Most
adaptations at the lower-level reactive layers in contrast are managed
by specialized components (for example LAAS architecture
\cite{alami:1998p820,Ingrand07} and CLARATy \cite{Nesnas:2003do}
) that rely on different representations for modeling local
control behavior or worse do not have any explicit formal agent
behavior model. The consequent different techniques for specifying
each layer in the architecture results in duplication of effort and a
diffusion of knowledge leading to significant design and integration
issues \cite{DS1report}.  \texttt{IDEA} mitigated most of these issues
by tightly integrating planning and execution in a single
representational and computational framework with a unified
declarative model. Our approach with \texttt{T-REX} builds on this
design methodology albeit in a substantially systematic manner

The \rx architecture contributes to agent architectures over and
beyond \texttt{IDEA} are in the following ways:

@@

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as
  each reactor declares its planning look-ahead $\pi$ and expected
  latency $\lambda$ before being able to produce plan to the agent
  depending on where it is located in a reactor dependency graph such
  as shown in Fig. \ref{fig:agent}.

\item \emph{Coupling} a tighter integration between the planning
  process and world evolution to ensure that planning occurs while
  being continuously informed of state change. By doing so, the
  planner is better informed in its deliberation despite a larger
  latency than changing world change state.

@@

Partitioned inference is a major contribution of \rx. However that is
not the focus of this chapter; details are in \cite{py10} and
\cite{rajan12}. Coupled interaction between planning and the world on
the other hand, has a significant implication on how planning can be
integrated in such frameworks; here we show using the \eu planner can
be coupled to our framework. First we link concepts in \eu planning
with the \rx framework. Note that in the following treatise, we call
the formal semantic framework of \rx as the 'agent'. Reactors
constitute an entity within this agent and are standalone blocks of
inference.

@@

The \rx architecture is structured around the notion of \eu {\em state
  variables} as the fundamental basis of interaction. Additionally it
comes with a well-defined ownership model associated with each state
variable. In this framework each reactor is seen as a 'black box'
providing its own {\em internal} state variables which it is
responsible for in order to maintain its value at every instant of
time. Each such instance is represented by a {\em tick} with a fixed
duration; in our implementation of \rx on the Dorado, we use a tick of
$1$ second. In order to identify its state a reactor can subscribe to
{\em external} state variables owned by other reactors, for which it
will receive new updates when the owner reactor generates its {\em
  internal} state.

As the execution frontier -- called $\tau$ -- advance, the \rx agent
propagates the current state information accross all the reactors states
variables through the {\em synchronization} process.
 The sequencing of the reactors synchronization is identified
by traversing the reactor graph from the least dependents reactors --
the ones with no {\em external} timelines -- to the most dependent
ones which typicallu do not provide {\em internal} timelines.

Conversely each reactor can request a future change on one of its {\em
  external} state variables which will be transformed as an {\em
  internal} goal to the owner.  Such \emph{goal dispatching} is
expected to produce a deliberation phase for the receiving reactor
which can cyclically and in turn, produce a plan impacting its {\em
  external} state variables' future that will propagate down following
the same mechanism.

In order to avoid to be overflood by goals that may threats the
ability of a reactor to process these goals and integrate them in tis
plan. In order to do so each reactor specifies the two following
values that apply for the full scope of the agent lifetime:
\begin{itemize}
\item $\lambda$ is the specified latency of this reactor, which is an
  indicator of the maximum expected time for a reactor to produce a
  plan.\footnote{Note that while it is highly recommended to select a
    sound value, a failure to produce a plan within a chosen value for
    this parameter is not considered as a critical failure of the
    reactor.}
\item $\pi$ is the planning look-ahead of the reactor. which indicates
  how many ticks in the future this reactor is looking ahead during
  deliberation. 
\end{itemize}

Using this information along with the dependency relation betwen
reactor \rx defines the notion of execution latency of a state
variable as follow:
\begin{definition}
  The {\em execution latency}  of a {\em state variable} $s$ which is
    internal to the reactor $r$ is the maximum expected delay before 
    a goal on this state varabale can be included in its plan and
    executed by the  reactor $r$.

    It is defined recursrevely as follow: 
    \begin{equation*}
      \lambda_{exec}(s) = \lambda(r) + \max_{e \in exts(r)}
      \lambda_{exec}(e) 
  \end{equation*}
    Where  $exts(r)$ is the set of the state varaibles which are
    {\em externals} to $r$. 
\end{definition}

The concept of execution latency is critical at the agent level
and can also impact the planning/deliberation process on the {\em
  external} state variables of a reactor. For simplicity sake in the
future sections more focused in the \eu reactor, we abstract it out 
as being equivalent to the reactor latency (meaning that we 
assume that all {\em external} timelines of this reactor have
$\lambda_{exec} = 0$). At the \rx agent level though this
 execution latency is used in turn to define the {\em dispatch
window} of the same state variable:
\comment{Needed to say that early on but the flow of this paragraph 
is not very tight}
\begin{definition}
  \label{def:dispatch}
  For every timelines $s$ {\em internals} to the reactor $r$, the {\em 
    dispatch window} of this timeline at the execution frontier $\tau$
  is the time interval:
  \begin{equation*}
    \Pi(s) = [\tau + \lambda_{exec}(s), \tau+\lambda_{exec}(s)+\pi(r)]
  \end{equation*}
\end{definition}

This window is used by \rx in order to identify when to dispatch a
request to a {\em reactor} as a new goal. This happens only if the
request start time -- represented as the interval of possible start
instants -- intersects this window. Requests further in the future 
are buffered by the agent until this window advanced enough to 
overlap their start time. This gives a simple mechanism in order to
request a change from the {\em external} state variable of a a reactor 
while avoiding to dispatch too much goals at once to the reactor that
declare this same state variable as {\em internal}. The choice of both
the latency and look-ahead of a reactor can be delicate as one is
dependent on the other. Latency also highly dependent on exogenous
paramaters such as the computing power of the host system and the
number of reactors needing to deliberate at once. As of today these
should be evaluated carefully by the system designers and while
having an important impact in the efficacy of the agent to find and
execute a global strategy they are just indicative and a failure from
a reactor to produce a plan in its latency is not considered as a
critical failure from \rx point of view.

Requests from a reactor can  be {\em recalled} by this reactor when it
they are not part of its plan anymore. For example if this reactor
need to replan due to an unexpected event, it will recall ol the
requests it did propduce that have not been observed yet. These
recalls are notified to the reactor that received this request as a
goal in order to remove it from its current objectives.

@@

\begin{figure}[!htb]
  \centering
  \subfloat[\small A \rx agent is composed of multiple reactors or
    control loops (rounded boxes) which are connected through state
    variables provided by one reactor ({\color{blue}blue} solid line)
    with multiple possible clients ({\color{green}green} dashed
    lines)]{\includegraphics[scale=0.45]{figs/AUV-agent.pdf}
  \label{fig:agent}
  }
  \qquad
  \subfloat[\small State information flow between the
    \texttt{Pilot} and the \texttt{Vehicle} reactors.
    Observations transit as a bottom-up flow 
    and apply to the execution frontier $\tau$. 
    Goal information flow from the top to the bottom and relates to 
    desired future values of state variables.]{
      \includegraphics[width=0.4\columnwidth]{figs/TREX_flow}
      %\vskip-1cm
      \label{fig:trex:flow}
    }
    \caption{\rx connection of reactors through state variables based on the
    AUV agent we use at sea.}
\end{figure}

Consider for example the \rx agent instance we use on our AUV as shown
in Fig. \ref{fig:agent}. In this instance all the reactors (symbolized
by colored boxes) do not have an explicit connection from one to
another but instead rely on a publish/subscribe model on shared state
variables. For instance the \textsf{Pilot} reactor needs the {\em
  externally} managed state information from \textsf{Sensor\_Data} and
\textsf{Nav\_cmd} state variables in order to maintain the
\emph{internal} state variables \textsf{Waypoint} and
\textsf{Nav\_policy}. This relation applies in two ways:

@@

\item In order to identify its current {\em internal} state values the
  \textsf{Pilot} needs to know the current {\em external} state 
  information it relies on. This information is providede in turn by
  the reactors that provided these timelines as {\em internal}. In
  this specific case, the current \texttt{Waypoint} where the vehicle
  is heading can be evaluated based on the \texttt{Nav\_cmd} executed
  along with usefull \texttt{Sensor\_Data}, both being provided by the
  \texttt{Vehicle} reactor.

@@

Given the above example, the \rx architecture therefore helps abstract
reactor interaction by constraining them to state variables. Each
reactor is therefore agnostic where its {\em external} state
information is managed as long as this information is available and
properly maintained for managing {\em internal} state
information. This {\em internal} state in turn may be used by other
reactors. Such a formalism also constrains reactor interaction to
state information exchange which comes in two forms as illustrated in
Figure \ref{fig:trex:flow}:

@@

\item {\em Observation on current state values}: This information
  propagates \emph{up} in the reactor dependency graph at the
  synchronization phase and occurs at the beginning of every
  tick. During this phase the agent provides updates to each reactor
  on their {\em external} state variables for this tick so as they
  could compute their {\em internal} state variable values for this
  same tick. By doing so we propagate a consistent view of that state
  of the world throughout all the reactors as time
  advances. %\comment{could do with a figure here}

\item {\em Goal request on future state}: This information propagates
  \emph{down} in the reactor dependency graph. In order to satisfy own
  {\em internal} objectives, a reactor may have a plan that relies on
  future values of one of its {\em external} state. When such a plan
  is identified the agent ensures that this part of the plan is given
  as a goal to the owner of the given {\em external} state
  variable(s). %\comment{could do with a figure here}

>>>>

A prime motivation behind the \rx architecture was to design a
controller which could bring automated planning closer to low-level
control of the vehicle while maintaining its reactivity.

Traditionally planning has been deemed computationally expensive with
planner performance restricting robot reactivity
\cite{ghallab04,Dias:2003ua}. In control architectures where task
planning is embedded, planning consequently remains at an abstract
level with the assumption that it would otherwise impede system
reactiveness and that the environment can change at a faster rate than
the planner can plan for. In such a situation, the agent may thrash if
the internal state of the plan gets out of synch with the actual state
of the world. As a result the planning problem managed {\em in situ}
has traditionally remained fairly detached from low-level
control. Most adaptations at the lower-level reactive layers in
contrast are managed by specialized components (for example
\texttt{LAAS architecture} \cite{alami:1998p820,Ingrand07} and
\texttt{CLARATy} \cite{Nesnas:2003do}) that rely on different
representations for modeling local control behavior or worse do not
have any explicit formal agent behavior model. The consequent
different techniques for specifying each layer in the architecture
results in duplication of effort and a diffusion of knowledge leading
to significant design and integration issues \cite{DS1report}.
\texttt{IDEA} mitigated most of these issues by tightly integrating
planning and execution in a single representational and computational
framework with a unified declarative model. Our approach with \rx
builds on this design methodology albeit in a substantially systematic
manner. The \rx architecture contributes to agent architectures over
and beyond \texttt{IDEA} are in the following ways:

@@

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as
  each reactor declares its planning look-ahead $\pi$ and expected
  latency $\lambda$ before being able to produce plan to the agent
  depending on where it is located in a reactor dependency graph such
  as shown in Fig. \ref{fig:agent}. A reactor closer to the hardware
  in the hierarchy, is expected to deliberate less and be highly
  responsive to observations from the world. This implies lower values
  of $\pi$ and $\lambda$ (see Section \ref{sec:arch:trex} below).

\item \emph{Coupling} a tighter integration between the planning
  process and world evolution to ensure that planning occurs
  \emph{during continuous} state update. By doing so, the planner is
  better informed in its deliberation despite a larger latency than
  the changing world.

@@

Systematic partitioned inference is a major contribution of
\rx. However that is not the focus of this chapter with details
covered in \cite{py10, rajan12}. Coupled interaction between planning
and the world on the other hand, has a significant implication on how
planning can be integrated in such frameworks. In this chapter we show
how using the \eu planner can be coupled to our framework by first
linking concepts in temporal planning with the \rx framework. Note
that in the following treatise, we call the formal semantic framework
of \rx as the 'agent'. Reactors constitute an entity within this agent
and are standalone blocks of inference. The scope, detail and number
of reactors are application specific and outside the scope of this
chapter.

@@

The \rx architecture is structured around the notion of \eu {\em state
  variables} as the fundamental basis of interaction. Additionally it
comes with a well-defined ownership model associated with each such
state variable. In this framework each reactor is seen as a 'black
box' providing its own {\em internal} state variable which it is
responsible for in order to maintain its value at every instant of
time. Each such instance is represented by a {\em tick} with a fixed
duration; in our implementation of \rx on the Dorado AUV, we use a
tick of $1$ second. In order to identify its state a reactor can
subscribe to {\em external} state variables owned by other reactors,
for which it will receive new updates when the owner reactor generates
its {\em internal} state.

\begin{figure}[!htbp]
  \centering
  \includegraphics[scale=0.35]{figs/DAG-hierarchy}
  \caption{\small An example of a \rx reactor hierarchy which is used
    to determine a synchronization order for propagation of values in
    \emph{external} state variables.}
  \label{fig:reactor-DAG}
\end{figure}

As the execution frontier, $\tau$ advances, the \rx agent propagates
the current state information across all reactor state variables
through the {\em synchronization} process.  The sequencing of the
reactors synchronization is identified by traversing a reactor
dependency graph from the least dependent -- ones with no {\em
  external} timelines -- to the most, \ie which typically do not have
{\em internal} timelines. Fig. \ref{fig:reactor-DAG} shows an example
of such a dependency graph implemented as a directed acyclic graph.

Conversely each reactor can request a future change on one of its {\em
  external} state variables which will be transformed as an {\em
  internal} goal to the owner reactor(s).  Such \emph{goal
  dispatching} is expected to produce a deliberation phase for the
receiving reactor which can cyclically in turn, produce a plan
impacting its {\em external} state variables' future that will
propagate down following the same mechanism.

In such a process, one challenge is to avoid being overwhelmed by
goals that may threaten the ability of a reactor to process and
integrate in its evolving partial plan. For that reason, each reactor
specifies the following two parameters that applied towards the scope
of the agent lifetime:

\item $\lambda$ is the specified latency of this reactor, which is an
  indicator of the \emph{maximum expected time} for a reactor to
  produce a plan.\footnote{Note that while it is highly recommended to
    select a sound value, a failure to produce a plan within a chosen
    value for this parameter is not considered as a critical failure
    of the reactor.}

\item $\pi$ is the planning \emph{look-ahead} of the reactor. which is
  indicative of the number of ticks in the future this reactor is
  looking ahead during deliberation.

Using this information along with the dependency relation between
reactors, \rx defines the notion of execution latency of a state
variable as follow:

\begin{definition}
  The {\em execution latency} of a {\em state variable} $s$ which is
  internal to the reactor $r$ is the maximum expected delay before a
  goal on this state variable can be included in its plan and executed
  by the reactor $r$. It is defined recursively as: 
  \begin{equation*}
    \lambda_{exec}(s) = \lambda(r) + \max_{e \in exts(r)}
    \lambda_{exec}(e) 
  \end{equation*}
  Where  $exts(r)$ is the set of the state variables which are
  {\em externals} to $r$. 
\end{definition}

For intertwined plan execution, the concept of execution latency is
critical at the agent level since it can impact the
planning/deliberation process on {\em external} state variables of a
reactor. For simplicity we abstract it out as being equivalent to the
reactor latency implying that we assume that all {\em external}
timelines of this reactor have $\lambda_{exec} = 0$. At the \rx agent
level this execution latency is used to define the {\em dispatch
  window} of the same state variable.% : \comment{Needed to say that
  % early on but the flow of this paragraph is not very tight}
\kcomment{Looks fine to me unless you want to clarify further.}

\begin{definition}
  \label{def:dispatch}
  For every timeline $s$ {\em internal} to a reactor $r$, the {\em
    dispatch window} of this timeline at the execution frontier $\tau$
  is the time interval:
  \begin{equation*}
    \Pi(s) = [\tau + \lambda_{exec}(s), \tau+\lambda_{exec}(s)+\pi(r)]
  \end{equation*}
\end{definition}

This window is used by \rx to identify when to dispatch a request to a
{\em reactor} as a new goal. This happens only if the request start
time -- represented as the interval of possible start times --
intersects this window. Requests further in the future are buffered by
the agent until this time window advances for an overlap. This
provides an elegant mechanism to request a change from the {\em
  external} state variable of a reactor while avoiding flooding the
owner reactor of this state variable.

The choice of both the latency and look-ahead of a reactor can be
delicate as one is dependent on the other. Needless to say, latency is
highly dependent on exogenous paramaters such as the host computers
computing power and the number of reactors needing to deliberate at
once, among others. As the design stands, these variables should be
evaluated carefully by the system designer balancing agent efficacy.

% efficacy of the agent to find and execute a
% global strategy they are just indicative and a failure from a reactor
% to produce a plan in its latency is not considered as a critical
% failure from \rx point of view.

Requests from a reactor can be {\em recalled} when it is determined
that they are not part of its projected plan anymore. For example if a
reactor needs to replan due to an unexpected event, it will recall all
requests it produced that have not been observed \kcomment{replace
  ``observed'' with ``not incorporated by the owner reactor''?}
yet. This recall is sent to the reactor that received the goal(s) to
remove it/them from its current set of objectives.

@@

\begin{figure}[!htb]
  \centering
  \subfloat[\small A \rx agent is composed of multiple reactors or
    control loops (rounded boxes) which are connected through state
    variables provided by one reactor ({\color{blue}blue} solid line)
    with multiple possible clients ({\color{green}green} dashed
    lines)]{\includegraphics[scale=0.45]{figs/AUV-agent.pdf}
  \label{fig:agent}
  }
  \qquad
  \subfloat[\small State information flow between the \texttt{Pilot}
  and the \texttt{Vehicle} reactors in an \rx instance.  Observations
  transit as a bottom-up flow and apply to the execution frontier
  $\tau$.  Goal information flow from the top to the bottom and
  relates to desired future values of state variables.]{
    \includegraphics[width=0.4\columnwidth]{figs/TREX_flow}
    \label{fig:trex:flow}
  }
\caption{\small \rx reactors and shared state variables.}
\end{figure}

Consider for example the \rx agent instance we use on our AUV as shown
in Fig. \ref{fig:agent}. In this instance all the reactors (symbolized
by colored boxes) % do not have an explicit connection from one to
% another but instead
rely on a publish/subscribe model on shared state variables. For
instance the \textsf{Pilot} reactor needs the {\em externally} managed
state information from \textsf{Sensor\_Data} and \textsf{Nav\_cmd}
state variables in order to maintain the \emph{internal} state
variables \textsf{Waypoint} and \textsf{Nav\_policy}. This relation
applies in two ways:

@@

\item In order to identify its current {\em internal} state values the
  \textsf{Pilot} needs to know the current {\em external} state
  information it relies on. This information is provided in turn by
  the reactors that define these timelines as {\em internal}. In this
  specific case, the current \texttt{Waypoint} where the vehicle is
  heading can be evaluated based on the \texttt{Nav\_cmd} executed
  along with useful \texttt{Sensor\_Data}, both being provided by the
  \texttt{Vehicle} reactor.

@@

Given the above example, \rx abstracts reactor interaction by
constraining them to state variables. Each reactor is therefore
agnostic where its {\em external} state information is managed as long
as this information is available and properly maintained for managing
{\em internal} state information. This {\em internal} state in turn
may be used by other reactors. Such a formalism also constrains
reactor interaction to state information exchange which comes in two
forms as illustrated in Fig. \ref{fig:trex:flow}:

@@

\item {\em Observation on current state values}: This information
  propagates \emph{up} in the reactor dependency graph at the
  synchronization phase and occurs at the beginning of every
  tick. During this phase the agent provides updates to each reactor
  on their {\em external} state variables for this tick so they can
  compute their {\em internal} state variable values for this same
  tick. By doing so we propagate a consistent view of that state of
  the world throughout all the reactors as time
  advances.

\item {\em Goal request on future state}: This information propagates
  \emph{down} in the reactor dependency graph. In order to satisfy own
  {\em internal} objectives, a reactor may have a plan that relies on
  future values of one of its {\em external} state. When such a plan
  is identified the agent ensures that this part of the plan is
  provided as a goal to the owner of the given {\em external} state
  variable(s).

### 2012-02-17 concepts.tex
europ+concepts touch ups with Javier, fig for arch and updates to arch

  \eu's ability to represent and reason over numeric intervals is a
  very useful characteristic that allow users to create flexible
  plans. \comment{say more about the benefits of flexibility?}

@@

\item[\textbf{Tokens}] In first order logic, a predicate defines a
  relation between objects and properties \comment{citation?}.  In
  \eu, we define such relations between variables whose domains are
  sets of objects and sets of properties to describe state and
  behavior. For example, we might use a predicate \texttt{At($a$,$l$)}
  to indicate that agent $a$ is at location $l$, or a predicate
  \texttt{Buy($a$,$p$}) to indicate that agent $a$ is taking acting to
  buy product $p$. Note that $a$ is a variable which may have a number
  of possible values in a problem with multiple agents. Similarly, $l$
  and $p$ are variables whose values are the set of possible positions
  and products respectively. \eu can be used to create partial plans,
  where the domains of variables $a$ and $l$ can have more than one
  possible value in them, or grounded plans, where single values will
  be specified for each variable as we saw in the case of the
  N-queen's problem.

@@

\item \textit{Timelines}: Often objects in a domain must be described
  by exactly one token for every given timepoint in the plan. Any
  instances of a class derived from \eu's built-in Timeline class will
  induce ordering requirements among its tokens to ensure no temporal
  overlap may occur among them \cite{mus94}.  \comment{add small
    example?}

\item \textit{Resources}: Metric resources, e.g. the energy of a
  battery or the capacity of a cargo hold, are objects with an
  explicit quantitative state in time and with a circumscribed range
  of changes that can occur to impact that state i.e. produce,
  consume, use, change.  Resources are such a common requirement for
  \eu users that built-in object types (with their corresponding token
  types to denote production, consumption, etc) are provided for
  them. Instances of classes derived from a Resource will induce
  ordering requirements on their Tokens in order to ensure that the
  level of the resource remains within specified limits.  \comment{add
    small example?}
\end{enumerate}

@@

  \comment{Add figure showing an AgentLocation timeline with an
    At(Home) token with start=0 and end open, an empty AgentBag
    timeline, and a Own(Drill) token with start=10 floating below both
    timelines}

@@

\comment{Add figure showing the effect of merging}

@@

\comment{Add figure showing the effect of activating the token by
  putting it on the AgentBag timeline, and a new At(Drill.location)
  token floating below}

@@

\comment{Add figure showing possible state transitions}

@@

  \comment{ Add figure showing internal and external relationships
    from rules on action Go}

@@

\comment {talk about condition/effect annotations?}

>>>>

  \eu's ability to represent and reason over numeric intervals is a
  very useful characteristic that allow users to create flexible
  plans.

@@

\item[\textbf{Tokens}] In first order logic, a predicate defines a
  relation between objects and properties \comment{use Genesereth citation?}.  In
  \eu, we define such relations between variables whose domains are
  sets of objects and sets of properties to describe state and
  behavior. For example, we might use a predicate \texttt{At($a$,$l$)}
  to indicate that agent $a$ is at location $l$, or a predicate
  \texttt{Buy($a$,$p$}) to indicate that agent $a$ is taking acting to
  buy product $p$. Note that $a$ is a variable which may have a number
  of possible values in a problem with multiple agents. Similarly, $l$
  and $p$ are variables whose values are the set of possible positions
  and products respectively. \eu can be used to create partial plans,
  where the domains of variables $a$ and $l$ can have more than one
  possible value in them, or grounded plans, where single values will
  be specified for each variable as we saw in the case of the
  N-queen's problem.

@@

\item \textit{Timelines}: Often objects in a domain must be described
  by exactly one token for every given timepoint in the plan. Any
  instances of a class derived from \eu's built-in Timeline class will
  induce ordering requirements among its tokens to ensure no temporal
  overlap may occur among them \cite{mus94}.  \comment{Maybe: add small
    example?}

\item \textit{Resources}: Metric resources, e.g. the energy of a
  battery or the capacity of a cargo hold, are objects with an
  explicit quantitative state in time and with a circumscribed range
  of changes that can occur to impact that state i.e. produce,
  consume, use, change.  Resources are such a common requirement for
  \eu users that built-in object types (with their corresponding token
  types to denote production, consumption, etc) are provided for
  them. Instances of classes derived from a Resource will induce
  ordering requirements on their Tokens in order to ensure that the
  level of the resource remains within specified limits.
  \comment{Maybe: add small example?}
\end{enumerate}

@@

  \comment{Use Fig 4 and edit:: Add figure showing an AgentLocation timeline with an
    At(Home) token with start=0 and end open, an empty AgentBag
    timeline, and a Own(Drill) token with start=10 floating below both
    timelines}

@@

\comment{Use previous fig and modify as appropriately:: Add figure
  showing the effect of merging; all figs on this section are
  conjoined. JB to sketch it out.}

@@

\comment{Use previous fig and modify as appropriately: Add figure
  showing the effect of activating the token by putting it on the
  AgentBag timeline, and a new At(Drill.location) token floating
  below}

@@

\comment{Maybe: Add figure showing possible state transitions. Need a FSM.}

@@

  \comment{Use previous fig and modify as appropriately (used as
    above): Add figure showing internal and external relationships
    from rules on action Go. Need a fig from JB}

@@
